---
title: "Mini-Project 02"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_float: true
---

# Data Visualization Project 02


```{r}
library(tidyverse)
library(plotly)
library(sf)
library(dplyr)
library(RColorBrewer)
library(ggplot2)
```

```{r}
getwd()
```

```{r}
data <- read_csv("../data/20_Victims_of_rape.csv")
data
```
# Interactive Plot


```{r}
inc_victims <- subset(data, Subgroup == "Victims of Incest Rape")

g <- inc_victims %>%
  group_by(Year) %>%
  summarise(Cases_Reported = sum(Rape_Cases_Reported)) %>% 
  rename(Year = Year, Cases_Reported = Cases_Reported)

```

Used `subset()` to filter only one kind of subgroup,group them by year and sum the reported cases.


```{r}
 fig <- plot_ly(g, x = ~Year, y = ~Cases_Reported, type = 'scatter',
    hoverinfo = "text",
    hovertemplate = 'No.of Cases Reported: %{y} <extra></extra>',
    mode = 'lines+markers',
    marker = list(color='purple'),
               line = list(color = "blue")) %>%
  layout(xaxis = list(title = "Year"), yaxis = list(title = "Cases Reported"),title = "<b> Total number of Cases throughout the years </b>")

fig

```

Used `plotly` to increase the interactivity and `layout()` to improve the axis labels. 

### Let's save a standalone HTML file


```{r}
library(htmlwidgets)

htmlwidgets::saveWidget(
                widget = fig, #the plotly object
                file = "reported_cases_figure.html", #the path & file name
                selfcontained = TRUE #creates a single html file
                )
```
Saved the above plot using `htmlwidgets` and it in HTML format.

# Spatial Visualization

```{r}
g1 <- inc_victims %>%
  group_by(Area_Name) %>%
  summarise(Cases_Reported = sum(Rape_Cases_Reported)) %>%
  rename(`State/UT` = Area_Name)
```

```{r}
shp_gdf <- st_read('../data/crime/Indian_states.shp')
```

Used `st_read()` to read the shape file.

```{r}
colnames(shp_gdf)
```


```{r}
merged <- left_join(shp_gdf, g1, by = c('st_nm' = 'State/UT'))
```

We merged the shape file and filtered data(by Area_Name) using `left_join()`.

```{r}
ggplot() +
  geom_sf(data = merged, aes(fill = Cases_Reported), color = 'black', lwd = 0.5) +
  scale_fill_gradient(name = 'Cases Reported', low = 'white', high = 'red') +
  coord_sf(crs = "+proj=merc") +
  labs(title = 'State-wise Rape-Cases Reported (2001-2010)',x = NULL, y = NULL) +
  scale_x_continuous(labels = function(x) paste0(x, '\u00B0', "E")) +
  scale_y_continuous(labels = function(x) paste0(x, '\u00B0', "N")) +
  theme_minimal() +
  theme(legend.position = 'bottom')

```

# Linear Model

```{r}
head(g,10)
```



```{r}
library(broom)

```

```{r}
linear_reg <- lm(Cases_Reported~Year, data=g)

summary(linear_reg)
```

```{r}
tidy(linear_reg)
```

`tidy()` constructs a tibble that summarizes the model’s statistical findings. This includes coefficients and p-values for each term in a regression

```{r}
glance(linear_reg)

```

```{r}
library(scales)
```


```{r}
ggplot(g, aes(x = Year, y = Cases_Reported)) +
 geom_point(color="#D55E00", size = 2) +
 geom_smooth(method = "lm", 
 formula = "y ~ x",
 color = "#009E73",
 size = 1.2,
 fill="yellow") + 
 scale_x_continuous(breaks = pretty_breaks()) +
 annotate("text", x = 2005, y = 700, label = "outlier") +
 theme_bw() +
 labs (x="Year", y="Cases Reported",
       title = "Linear Regression Model",
       caption ="Dataset : rape_cases") 
```

Plot the linear regression model using `geom_point()` and `geom_smooth()` along with `annotate()` to display the outlier.

